Family Fortunes: A Quantitative Analysis of the Early Iron Age Cemeteries at Knossos, Crete

### 2020-10-13 analysis/paper/Family_Fortunes_Article.Rmd
Repo name changes

```{r map, fig.align= "centre", fig.cap="goes around here"}

@@

```{r allpithoi,fig.dim = c(8, 3), dpi=600, fig.align='center', fig.cap="goes around here"}

@@

```{r aoristbycem, dpi=600, fig.align="center", fig.cap="goesaround here.", fig.dim=c(8,3.5)}
KNC<-subset(tombs,tombs$cem_condensed=="KNC")

@@

```{r tombconstruction, fig.dim=c(7.5,4), dpi=600, fig.cap="goes around here"}

@@

```{r tombsinuse, fig.dim=c(3,3),out.width="25%", dpi=600, fig.cap="goes around here"}

@@

ggplot(tombs_in_use_cem,aes(period,no_tombs, fill=cemetery))+geom_col(position="dodge")+theme_classic()+
       theme_bw()+theme(legend.title=element_blank(),
          text=element_text(family="serif", size=10),legend.position = "bottom")+
            labs(x="Period", y="Tombs in Use")+
              scale_fill_manual(values =c("gray50","gray80"),labels=c(" KNC  "," Fortetsa"))+ 
                 scale_y_continuous(expand=c(0,0),limits=c(0,45))
```

@@

```{r tombsinusebybuilddate, fig.dim=c(7,4), dpi=600, fig.cap="goes around here"}
tombs_11th<-subset(tombs,tombs$century_built=="11th")
tombs_10th<-subset(tombs,tombs$century_built=="10th")
tombs_9th<-subset(tombs,tombs$century_built=="9th")
tombs_8th<-subset(tombs,tombs$century_built=="8th")
tombs_7th<-subset(tombs,tombs$century_built=="7th")

@@

```{r potsbydateandtombdate,  fig.dim=c(3,7),out.width="0.5%", dpi=600, fig.cap="goes around here"}

@@

```{r toptombs, fig.dim = c(7, 5), dpi=600, fig.align='center',fig.cap="goes around here"}

@@

ggplot(aorist_all_tomb_melt,aes(variable, value,fill=cemetery))+geom_col()+theme(legend.position = "none")+facet_wrap(~tomb,scales="free_y")+
  geom_col()+theme_bw()+theme(legend.title=element_blank(),legend.position="bottom",
      text=element_text(family="serif", size=12),
      axis.text.x=element_text(angle=45, vjust=1.2, hjust=1))+
          labs(x="Date (Years BC)", y="Aoristic Sums")+
              scale_fill_manual(values = c("gray50","gray80"),labels=c(" KNC  "," Fortetsa"))+ 
                  coord_cartesian(xlim=c(1100,610))+scale_y_continuous()+
                    scale_x_reverse(expand = c(0, 0),limits=c(1100,610))+
                      theme(plot.margin=unit(c(0.1,1,0.1,0.1),"cm"))

@@

```{r groupsab, fig.dim=c(8,4),dpi=600, fig.align='center',fig.cap="goes around here"}

@@

```{r dimensionsboxplots, fig.dim=c(7,7),dpi=600, fig.align='center',fig.cap="goes around here"}

@@

```{r tombgroupdimensions, fig.dim = c(3.5, 3.5),out.width="10%", dpi=600, fig.align='center', fig.cap="goes around here"}

ggplot(subset(tombs,century_built=="9th" | century_built=="10th"),aes(dromos_length,chamber_area,color=group,label=tomb))+geom_point(size=1)+
  theme_bw()+theme(text=element_text(family="serif", size=10),legend.position = "bottom", legend.margin=margin(0,0,0,0),legend.box.margin=margin(-10,-10,-10,-10))+
  geom_label_repel(data=subset(tombs,group!="C"),aes(fill=group),color="white",show.legend = FALSE,family="serif")+
            labs(x="Dromos Length (m)", y="Chamber Area (sq m)")+
               scale_color_brewer(palette = "Set2",name="Group")+
               scale_fill_brewer(palette = "Set2")+
                scale_x_continuous(expand=c(0,0),limits=c(0,8))+
                  scale_y_continuous(expand=c(0,0),limits=c(0,10))+
                      theme(plot.margin=unit(c(0.1,1,0.1,0.1),"cm"))

@@

```{r tombdimensions, fig.dim = c(7.5, 4), dpi=600, fig.align='center', fig.cap="goes around here"}

@@

```{r tombscompthroughtime, fig.dim=c(7,4),dpi=600,fig.align='center',fig.cap="goes around here"}

@@

ggplot(pots9th10th,aes(tomb,fill=group))+geom_bar()+
  theme_bw()+theme(legend.position="bottom",text=element_text(family="serif", size=10),
               axis.text.x = element_text(angle = 90,hjust=1,vjust=0.4))+
                 guides(fill=guide_legend(title="Group"))+labs(x="Tomb", y="Vessel Count")+
                  scale_fill_brewer(palette = "Set2")+ 
                    scale_y_continuous(expand=c(0,0))+facet_grid(~century,scales="free_x",space="free")

@@

```{r bigbox, fig.dim=c(7,4),dpi=600, fig.align='center',fig.cap="goes around here"}

@@

ggplot(aorist_melt,aes(variable,value,fill=period))+geom_boxplot(aes(group = cut_width(variable, 10)),outlier.size=1,lwd=0.3)+theme_bw()+
      theme(text=element_text(family="serif", size=10))+
            labs(x="Date (Years BC)", y="Aoristic Sums per Tomb")+
               scale_fill_brewer(name="Period",palette = "Set3")+
                scale_x_reverse(expand=c(0,0),limits=c(1000,600))+
                  scale_y_continuous(expand=c(0,0),limits=c(0,20))

@@

```{r gini, fig.dim=c(7,4),dpi=600, fig.align='center',fig.cap="goes around here"}

@@

```{r imports, fig.dim=c(7,4),dpi=600, fig.align='center',fig.cap="goes around here"}

@@

ggplot()+ geom_col(data=all_import_imitation,aes(decade,aoristic_sum, fill=origin),position="dodge")+
            theme_bw()+theme(text=element_text(family="serif", size=10), legend.position="bottom")+
                labs(x="Date (Years BC)", y="Aoristic Sum")+
                  scale_fill_brewer(name="Origin",palette = "Set2")+
                   scale_y_continuous(expand=c(0,0),limits=c(0,20))+
                    scale_x_reverse(expand = c(0, 0),limits=c(1000,600))

@@

```{r comps, fig.dim=c(7,4),dpi=600, fig.align='center',fig.cap="goes around here"}

@@

potcompKNC<-ggplot(subset(pots,century!="" & cem_condensed=="NC"),aes(century,fill=type))+geom_bar(position="fill")+
  theme_bw()+theme(plot.title = element_text(hjust=0.5),legend.title = element_blank(),
               text=element_text(family="serif", size=12))+
                 labs(title="KNC",x="", y="Proportion of Total Assemblage")+
                  scale_fill_brewer(palette = "Set2")+ 
                    scale_y_continuous(expand=c(0,0.0))
potcompFOR<-ggplot(subset(pots,century!="" & cem_condensed=="FOR"),aes(century,fill=type))+geom_bar(position="fill")+
  theme_bw()+theme(plot.title = element_text(hjust=0.5),legend.title=element_blank(),
               text=element_text(family="serif", size=12))+
                 labs(title="Fortetsa",x="Century BC", y="")+
                  scale_fill_brewer(palette = "Set2")+ 
                    scale_y_continuous(expand=c(0,0.0))

potcomptop7<-ggplot(subset(pots,century!="" & century!="11th"& group=="A"),aes(century,fill=type))+geom_bar(position="fill")+
  theme_bw()+theme(plot.title = element_text(hjust=0.5),legend.title=element_blank(),
               text=element_text(family="serif", size=12))+
                 labs(title="Group A",x="", y="")+
                  scale_fill_brewer(palette = "Set2")+ 
                    scale_y_continuous(expand=c(0,0.0))

@@

```{r hists, fig.dim=c(6,4),dpi=600, fig.align='center',fig.cap="goes around here"}
tombs$group<-revalue(tombs$group,c("A"="A","B"="B","C"="B"))
total_object<-ggplot(subset(tombs,est_burials>1 & total_object>0),aes(total_object, fill=group))+geom_histogram()+
  theme_bw()+theme(legend.title=element_blank(),
               text=element_text(family="serif", size=10))+
                 labs(x="Total Non-Ceramic Finds", y="Number of Tombs")+
                  scale_fill_manual(values = c("gray80","gray50"),labels=c(" Group A  ", " Other Tombs"))+
                    scale_y_continuous(expand=c(0,0))+scale_x_continuous(expand=c(0,0))
total_weapon<-ggplot(subset(tombs,est_burials>1 & weapon>0),aes(weapon, fill=group))+geom_histogram()+
  theme_bw()+theme(legend.title=element_blank(),
               text=element_text(family="serif", size=10))+
                 labs(x="Total Weapons", y="Number of Tombs")+
                 scale_fill_manual(values = c("gray80","gray50"),labels=c(" Group A  ", " Other Tombs"))+
                    scale_y_continuous(expand=c(0,0))+scale_x_continuous(expand=c(0,0))
total_jewellery<-ggplot(subset(tombs,est_burials>1 & jewellery>0),aes(jewellery, fill=group))+geom_histogram()+
  theme_bw()+theme(legend.title=element_blank(),
               text=element_text(family="serif", size=10))+
                 labs(x="Total Jewellery", y="Number of Tombs")+
                  scale_fill_manual(values = c("gray80","gray50"),labels=c(" Group A  ", " Other Tombs"))+
                    scale_y_continuous(expand=c(0,0))+scale_x_continuous(expand=c(0,0))

@@

It remains to consider how we might characterise the burying groups behind these various, intriguing trends. The traditional view is that such groups were immediate families (Brock 1957, 41; @Boardman1967: 63; @Coldstream1996: 291). The relatively short lifespan of many tombs has been taken as evidence for direct, lineal inheritance, with the tomb going out of use as families moved, individuals founded new tombs, or the line fissioned or ended [@Cavanagh1996: 659]. The idea of family groups is supported by the, admittedly circumstantial, evidence of men, women and children being found within single tombs, and one instance of a (putatively) hereditary mandibular deformity in three skulls in the same tomb [@Musgrave1996].

@@

##### pagebreak
# Captions

*Figure 1 Map of the Knossos area, showing the locations of the various EIA tomb clusters and cemeteries, with EIA pottery distributions from the Knossos Urban Landscape Project marked in orange, giving an impression of the contemporary urban extent (underlying image credit: T. Whitelaw).*

*Figure 2 Relationship between known interments, counts of pithoi, and counts of pithoi, amphorae and kraters across both cemeteries. The straight line represents the mean estimate based on these three values.*

*Figure 3 Counts of pithoi across both cemeteries through time. Vessel dates have here been based on the midpoint (or mean) of the absolute date range accompanying their respective ceramic periods.*

*Figure 4 Aoristic sums of all vessels across both cemeteries. This represents a probabilistically weighted visualisation of depositional activity through time.*

*Figure 5 Aoristic sums of pithoi (left), as a proxy for cremation burials, and of all vessels (right). The sums are colour coded to show the values of the two cemeteries, and the total values*

*Figure 6 Tomb construction through time at both cemeteries. On the left, counts are based on the ceramic phase considered most likely to correspond to the tomb's first use (including simply 'PG', 'G' and 'O' where no greater specificity was possible). On the right, aoristic sums are presented, spreading the uncertainty associated with those general assignations. In both cases, tombs lacking any securely dateable pottery are excluded.*
*Abbreviations*
*SM - Subminoan; (E/M/L)PG - (Early/Middle/Late) Protogeometric; PGB - Protogeometric B; (E/M/L)G - (Early/Middle/Late) Geometric; (E/L)O - (Early/Late) Orientalising*

*Figure 7 Estimates of the number of tombs in use during each of the ceramic phases attested in both cemetaries, based on counts of tombs yielding vessels of each period.*

*Figure 8 Estimates of the number of tombs in use through time across both cemeteries. The grey background represents the total number of tombs in use, while the smaller coloured bars correspond to tombs built in each of the five centuries of EIA activity at the cemeteries.*

*Figure 9 Counts of vessels from all tombs in both cemeteries, arranged, on the left, by the century to which they are assigned based on their ceramic period and, on the right, by the centuries in which the tombs they come from were built.*

*Figure 10 Aoristic sums for the 12 tombs with the largest ceramic assemblages, arranged in decreasing order. Note, for the sake of better illustrating the relative temporal patterns, rather than absolute values, the scales on each plot are distinct.*

*Figure 11 Aoristic sums for the Group A and Group B tombs, as defined above.*

*Figure 11 Boxplots comparing various tomb dimensions, divided, in each case, by the century of tomb construction.*

*Figure 12 Scatterplot comparing the dimensions of* dromos *length and chamber area. Groups A and B, defined earlier, are illustrated and labelled in green and orange respectively, while all remaining tombs ('Group C') are in blue.*

*Figure 13 Scatterplots comparing tomb dimensions (chamber area and* dromos *length) to the estimated total number of burials by tomb. The points area colour-coded as per Figure 12.*

*Figure 14 Counts of vessels dating to the 9th, 8th and 7th centuries respectively, found in tombs built in the 10th or 9th century which yielded upwards of 50 vessels when excavated. The tombs are colour-coded as per Figure 12.*

*Figure 15 Boxplot showing the aoristic sums of all vessels across both cemeteries through time, colour-coded by ceramic period. Boxes represent the interquartile range, horizontal lines the median value, whiskers values within 1.5 times the interquartile range, and points outliers beyond this. Note, the Y axis has been limited to 20 to aid readability, meaning the outliers Tomb G in the EG period, and Tomb P in the late EO-LO period are excluded, with values ranging from 29-34, and 36-44 respectively.*

*Figure 16 Gini coefficient values through time based on the aoristic sums for all vessels across both cemeteries. Background colours correspond to ceramic periods.*

*Figure 17 Aoristic sums of imported and 'oriental' imitation vessels through time across both cemeteries, colour-coded by regional origin in the case of imports.*

*Figure 18 The composition of the total pottery assemblage for each century across both cemeteries, and among the 7 notable tombs (Group A) drawn out earlier in the analysis.*

*Figure 19 Histograms corresponding, first, to counts of jewellery (beads, necklaces, rings, diadems etc.), metal objects, and imported items by tomb across both cemeteries and, secondly, to those counts divided by the estimated number of burials in each tomb.*

##### pagebreak
# Figures

##### pagebreak
# Tables

>>>>

```{r map, fig.align= "centre", fig.cap="Map of the Knossos area, showing the locations of the various EIA tomb clusters and cemeteries, with EIA pottery distributions from the Knossos Urban Landscape Project marked in orange, giving an impression of the contemporary urban extent (underlying image credit: T. Whitelaw)."}

@@

```{r allpithoi,fig.dim = c(8, 3), dpi=600, fig.align='center', fig.cap="(Left) Counts of pithoi across both cemeteries through time. Vessel dates have here been based on the midpoint (or mean) of the absolute date range accompanying their respective ceramic periods. (Right) Aoristic sums of all vessels across both cemeteries. This represents a probabilistically weighted visualisation of depositional activity through time."}

@@

```{r aoristbycem, dpi=600, fig.align="center", fig.cap="Aoristic sums of pithoi (left), as a proxy for cremation burials, and of all vessels (right). The sums are colour coded to show the values of the two cemeteries, and the total values.", fig.dim=c(8,3.5)}
KNC<-subset(tombs,tombs$cem_condensed=="KNC")

@@

```{r tombconstruction, fig.dim=c(7.5,4), dpi=600, fig.cap="Tomb construction through time at both cemeteries. On the left, counts are based on the ceramic phase considered most likely to correspond to the tomb's first use (including simply 'PG', 'G' and 'O' where no greater specificity was possible). On the right, aoristic sums are presented, spreading the uncertainty associated with those general assignations. In both cases, tombs lacking any securely dateable pottery are excluded. Abbreviations: SM - Subminoan; (E/M/L)PG - (Early/Middle/Late) Protogeometric; PGB - Protogeometric B; (E/M/L)G - (Early/Middle/Late) Geometric; (E/L)O - (Early/Late) Orientalising"}

@@

```{r tombsinuse, fig.dim=c(3,3),out.width="25%", dpi=600, fig.cap="Estimates of the number of tombs in use during each of the ceramic phases attested in both cemetaries, based on counts of tombs yielding vessels of each period."}

@@

ggplot(tombs_in_use_cem,aes(period,no_tombs, fill=cemetery))+geom_col(position="dodge")+theme_classic()+
       theme_bw()+theme(legend.title=element_blank(),
          text=element_text(family="serif", size=10),legend.position = "bottom", legend.margin=margin(0,0,0,0),legend.box.margin=margin(-10,-10,10,-10))+
            labs(x="Period", y="Tombs in Use")+
              scale_fill_manual(values =c("gray50","gray80"),labels=c(" KNC  "," Fortetsa"))+ 
                 scale_y_continuous(expand=c(0,0),limits=c(0,45))
```

@@

```{r tombsinusebybuilddate, fig.dim=c(7,4), dpi=600, fig.cap="Estimates of the number of tombs in use through time across both cemeteries. The grey background represents the total number of tombs in use, while the smaller coloured bars correspond to tombs built in each of the five centuries of EIA activity at the cemeteries."}
tombs_11th<-subset(tombs,tombs$century_built=="11th")
tombs_10th<-subset(tombs,tombs$century_built=="10th")
tombs_9th<-subset(tombs,tombs$century_built=="9th")
tombs_8th<-subset(tombs,tombs$century_built=="8th")
tombs_7th<-subset(tombs,tombs$century_built=="7th")

@@

```{r potsbydateandtombdate,  fig.dim=c(3,7),out.width="0.5%", dpi=600, fig.cap="Counts of vessels from all tombs in both cemeteries, arranged, above, by the century to which they are assigned based on their ceramic period and, below, by the centuries in which the tombs they come from were built."}

@@

```{r toptombs, fig.dim = c(7, 5), dpi=600, fig.align='center',fig.cap="Aoristic sums for the 12 tombs with the largest ceramic assemblages, arranged in decreasing order. Note, for the sake of better illustrating the relative temporal patterns, rather than absolute values, the scales on each plot are distinct."}

@@

ggplot(aorist_all_tomb_melt,aes(variable, value,fill=cemetery))+geom_col()+theme(legend.position = "none")+facet_wrap(~tomb,scales="free_y")+
  geom_col()+theme_bw()+theme(legend.title=element_blank(),legend.position="bottom",
      text=element_text(family="serif", size=12),
      axis.text.x=element_text(angle=45, vjust=1.2, hjust=1), legend.margin=margin(0,0,0,0),legend.box.margin=margin(-10,-10,10,-10))+
          labs(x="Date (Years BC)", y="Aoristic Sums")+
              scale_fill_manual(values = c("gray50","gray80"),labels=c(" KNC  "," Fortetsa"))+ 
                  coord_cartesian(xlim=c(1100,610))+scale_y_continuous()+
                    scale_x_reverse(expand = c(0, 0),limits=c(1100,610))+
                      theme(plot.margin=unit(c(0.1,1,0.1,0.1),"cm"))

@@

```{r groupsab, fig.dim=c(8,4),dpi=600, fig.align='center',fig.cap="Aoristic sums for the Group A and Group B tombs, as defined above."}

@@

```{r dimensionsboxplots, fig.dim=c(7,7),dpi=600, fig.align='center',fig.cap="Boxplots comparing various tomb dimensions, divided, in each case, by the century of tomb construction."}

@@

```{r tombgroupdimensions, fig.dim = c(3.5, 3.5),out.width="10%", dpi=600, fig.align='center', fig.cap="Scatterplot comparing the dimensions of* dromos *length and chamber area. Groups A and B, defined earlier, are illustrated and labelled in green and orange respectively, while all remaining tombs ('Group C') are in blue."}

ggplot(subset(tombs,century_built=="9th" | century_built=="10th"),aes(dromos_length,chamber_area,color=group,label=tomb))+geom_point(size=1)+
  theme_bw()+theme(text=element_text(family="serif", size=10),legend.position = "bottom", legend.margin=margin(0,0,0,0),legend.box.margin=margin(-10,-10,10,-10))+
  geom_label_repel(data=subset(tombs,group!="C"),aes(fill=group),color="white",show.legend = FALSE,family="serif", size=3)+
            labs(x="Dromos Length (m)", y="Chamber Area (sq m)")+
               scale_color_brewer(palette = "Set2",name="Group")+
               scale_fill_brewer(palette = "Set2")+
                scale_x_continuous(expand=c(0,0),limits=c(0,8))+
                  scale_y_continuous(expand=c(0,0),limits=c(0,10))+
                      theme(plot.margin=unit(c(0.1,1,0.1,0.1),"cm"))

@@

```{r tombdimensions, fig.dim = c(7.5, 4), dpi=600, fig.align='center', fig.cap="Scatterplots comparing tomb dimensions (chamber area and* dromos *length) to the estimated total number of burials by tomb. The points area colour-coded as per Figure 11."}

@@

```{r tombscompthroughtime, fig.dim=c(7,4),dpi=600,fig.align='center',fig.cap="Counts of vessels dating to the 9th, 8th and 7th centuries respectively, found in tombs built in the 10th or 9th century which yielded upwards of 50 vessels when excavated."}

@@

ggplot(pots9th10th,aes(tomb,fill=group))+geom_bar()+
  theme_bw()+theme(legend.position="bottom",text=element_text(family="serif", size=10),
               axis.text.x = element_text(angle = 90,hjust=1,vjust=0.4), legend.margin=margin(0,0,0,0),legend.box.margin=margin(-10,-10,10,-10))+
                 guides(fill=guide_legend(title="Group"))+labs(x="Tomb", y="Vessel Count")+
                  scale_fill_brewer(palette = "Set2")+ 
                    scale_y_continuous(expand=c(0,0))+facet_grid(~century,scales="free_x",space="free")

@@

```{r bigbox, fig.dim=c(7,4),dpi=600, fig.align='center',fig.cap="Boxplot showing the aoristic sums of all vessels across both cemeteries through time, colour-coded by ceramic period. Boxes represent the interquartile range, horizontal lines the median value, whiskers values within 1.5 times the interquartile range, and points outliers beyond this. Note, the Y axis has been limited to 20 to aid readability, meaning the outliers Tomb G in the EG period, and Tomb P in the late EO-LO period are excluded, with values ranging from 29-34, and 36-44 respectively."}

@@

ggplot(aorist_melt,aes(variable,value,fill=period))+geom_boxplot(aes(group = cut_width(variable, 10)),outlier.size=0.9,lwd=0.3)+theme_bw()+
      theme(text=element_text(family="serif", size=10))+
            labs(x="Date (Years BC)", y="Aoristic Sums per Tomb")+
               scale_fill_brewer(name="Period",palette = "Set3")+
                scale_x_reverse(expand=c(0,0),limits=c(1000,600))+
                  scale_y_continuous(expand=c(0,0),limits=c(0,20))

@@

```{r gini, fig.dim=c(7,4),dpi=600, fig.align='center',fig.cap="Gini coefficient values through time based on the aoristic sums for all vessels across both cemeteries. Background colours correspond to ceramic periods."}

@@

```{r imports, fig.dim=c(7,4),dpi=600, fig.align='center',fig.cap="Aoristic sums of imported and 'oriental' imitation vessels through time across both cemeteries, colour-coded by regional origin in the case of imports."}

@@

ggplot()+ geom_col(data=all_import_imitation,aes(decade,aoristic_sum, fill=origin),position="dodge")+
            theme_bw()+theme(text=element_text(family="serif", size=10), legend.position="bottom", legend.margin=margin(0,0,0,0),legend.box.margin=margin(2,-10,10,-10))+
                labs(x="Date (Years BC)", y="Aoristic Sum")+
                  scale_fill_brewer(name="Origin",palette = "Set2")+
                   scale_y_continuous(expand=c(0,0),limits=c(0,20))+
                    scale_x_reverse(expand = c(0, 0),limits=c(1000,600))

@@

```{r comps, fig.dim=c(7,4),dpi=600, fig.align='center',fig.cap="The composition of the total pottery assemblage for each century across both cemeteries, and among the 7 notable tombs (Group A) drawn out earlier in the analysis."}

@@

potcompKNC<-ggplot(subset(pots,century!="" & cem_condensed=="NC"),aes(century,fill=type))+geom_bar(position="fill")+
  theme_bw()+theme(plot.title = element_text(hjust=0.5),legend.title = element_blank(),
               text=element_text(family="serif", size=11))+
                 labs(title="KNC",x="", y="Proportion of Total Assemblage")+
                  scale_fill_brewer(palette = "Set2")+ 
                    scale_y_continuous(expand=c(0,0.0))
potcompFOR<-ggplot(subset(pots,century!="" & cem_condensed=="FOR"),aes(century,fill=type))+geom_bar(position="fill")+
  theme_bw()+theme(plot.title = element_text(hjust=0.5),legend.title=element_blank(),
               text=element_text(family="serif", size=11))+
                 labs(title="Fortetsa",x="Century BC", y="")+
                  scale_fill_brewer(palette = "Set2")+ 
                    scale_y_continuous(expand=c(0,0.0))

potcomptop7<-ggplot(subset(pots,century!="" & century!="11th"& group=="A"),aes(century,fill=type))+geom_bar(position="fill")+
  theme_bw()+theme(plot.title = element_text(hjust=0.5),legend.title=element_blank(),
               text=element_text(family="serif", size=11))+
                 labs(title="Group A",x="", y="")+
                  scale_fill_brewer(palette = "Set2")+ 
                    scale_y_continuous(expand=c(0,0.0))

@@

```{r hists, fig.dim=c(6,4),dpi=600, fig.align='center',fig.cap="Histograms corresponding, first, to counts of jewellery (beads, necklaces, rings, diadems etc.), metal objects, and imported items by tomb across both cemeteries and, secondly, to those counts divided by the estimated number of burials in each tomb."}

tombs$group<-revalue(tombs$group,c("A"="A","B"="B","C"="B"))
total_object<-ggplot(subset(tombs,est_burials>1 & total_object>0),aes(total_object, fill=group))+geom_histogram()+
  theme_bw()+theme(legend.title=element_blank(),
               text=element_text(family="serif", size=10))+
                 labs(x="Total Non-Ceramic Finds", y="Number of Tombs")+
                  scale_fill_manual(values = c("gray80","gray50"),labels=c(" Group A  ", " Other Tombs"))+
                    scale_y_continuous(expand=c(0,0))+scale_x_continuous(expand=c(0,0))
total_weapon<-ggplot(subset(tombs,est_burials>1 & weapon>0),aes(weapon, fill=group))+geom_histogram()+
  theme_bw()+theme(legend.title=element_blank(),
               text=element_text(family="serif", size=10))+
                 labs(x="Total Weapons", y="Number of Tombs")+
                 scale_fill_manual(values = c("gray80","gray50"),labels=c(" Group A  ", " Other Tombs"))+
                    scale_y_continuous(expand=c(0,0))+scale_x_continuous(expand=c(0,0))
total_jewellery<-ggplot(subset(tombs,est_burials>1 & jewellery>0),aes(jewellery, fill=group))+geom_histogram()+
  theme_bw()+theme(legend.title=element_blank(),
               text=element_text(family="serif", size=10))+
                 labs(x="Total Jewellery", y="Number of Tombs")+
                  scale_fill_manual(values = c("gray80","gray50"),labels=c(" Group A  ", " Other Tombs"))+
                    scale_y_continuous(expand=c(0,0))+scale_x_continuous(expand=c(0,0))

@@

It remains to consider how we might characterise the burying groups behind these various, intriguing trends. The traditional view is that such groups were immediate families [@Brock1957: 41; @Boardman1967: 63; @Coldstream1996: 291]. The relatively short lifespan of many tombs has been taken as evidence for direct, lineal inheritance, with the tomb going out of use as families moved, individuals founded new tombs, or the line fissioned or ended [@Cavanagh1996: 659]. The idea of family groups is supported by the, admittedly circumstantial, evidence of men, women and children being found within single tombs, and one instance of a (putatively) hereditary mandibular deformity in three skulls in the same tomb [@Musgrave1996].

@@
